Performance comparison of Dask and Apache Spark on HPC systems \HL{for Neuroimaging}

### 2022-09-09 latex/paper.tex
[TMP] [WIP] Current writing.

\HL{
	We configure Dask to have eight worker processes per node, each with eight threads.
	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
	Similarly, we configured Spark to have eight executors per node, each with eight threads.
	When trying different configuration for the workers, we did not observe performance difference
	for our benchmark.
	Therefore we chose the recommended configuration from Dask.
}
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
A new Dask or Spark cluster was started and torn down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

@@

\subsection{Kmeans}
Figure \ref{fig:kmeans_worker} depicts the total execution time for the \textit{Kmeans} application when varying the number of workers.
We observe that Dask is multifold faster than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
Specifically, Spark has much longer overhead, \textit{get\_labels} and read time than Dask.
This difference arises from Spark workers restarting due to lack of executor memory, leading to recomputation of tasks.

@@

\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_kmeans_worker.jpg}
	\caption{Kmeans: varying nodes -- 5000 files, 3 iterations}
	\label{fig:kmeans_worker}
\end{figure}

In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
of this new experiments for Dask and Spark.
The x-axis shows the time in second while the y-axis depicts the workers, with 
one worker thread ID (Dask) and process ID (Spark) per line.
We represent each task on the Gantt chart with its associated function's color.
We observe that Spark reports three times more processes than Dask threads.
\HL{
	It is unclear what is the cause for reporting extra Spark processes.
	However, with our analysis we found that both Dask and Spark use the same 
	amount of maximum concurrent processes or threads.
	Additionally, the is no significant difference in the average amount of 
	concurrent processes or threads between Dask and Spark.
}

@@

\subsection{Memory management}
Our Kmeans results from Figure~{\ref{fig:kmeans_worker}} show a significant difference between Dask and Spark.
The difference in performance comes from Spark worker restarting due to running out of memory; thus, Spark needs to recompute several tasks.
This is unexpected, given that Dask and Spark are configured similarly and are allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark using processes compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

\subsection{Recommendations}
\iffalse
The differences in performance are minimal between both engines. However, 
we added a “Recommendation” subsection to the discussion for qualitative 
difference and some caveats to avoid when using the engines.
\fi
\HL{
	The results from our experiments show that the performance difference between
	Dask and Spark is minimal, in general. 
	The main differences we observed are qualitative.
	We found that Dask integration with the scientific Python ecosystem made it
	easier to implement neuroimaging application.
	Spark configuration required to set multiple environment variables.
	In comparison, Dask is ready-to-go after an installation through PIP.
	Dask dashboard provides more information than Spark ones, making
	the debugging of Dask application easier and faster.
	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
	However, in other fields with non-Python-based ecosystem, both Dask and spark
	would be equal candidate as a Big Data engine.
}

\HL{
	% TODO add caveats to avoid when using the engines.
}

\subsection{Limitations}
To make a fair comparison between Spark and Dask, we set both engine configurations
to be equivalent to the best of our capabilities, which comes with some limitations.
Firstly, both engines use a shared implementation for the algorithms.
This is done to keep the theoretical runtime equivalent; however, the chosen
implementation might not be optimal for either of the engines.
Secondly, we decided to use Lustre as a file system for both Spark and Dask.
Spark could benefit from using HDFS as a file system, but HDFS is uncommon in HPC.
Finally, our infrastructure has a bottleneck from the network and Lustre.
As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer 
time is tightly correlated to the overhead of the application.
Therefore, while reducing this bottleneck should improve the performance of 
both engines similarly, it can be hard to predict the actual behavior.

>>>>

\HL{
	We configure Dask to have eight worker processes per node, each with eight threads.
	This is the default recommended configuration for the Dask SLURM cluster configuration~{\cite{DaskSLURMDoc}}.
	Similarly, we configured Spark to have eight executors per node (hereafter called workers for consistency with Dask), each with eight threads.
	When trying different configuration for the workers, we did not observe performance difference
	for our benchmark.
	Therefore we chose the recommended configuration from Dask.
}
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
A new Dask or Spark cluster was started and torn down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

@@

\subsection{Kmeans}
\HL{
	When varying the number of workers, we observed that Dask is multifold faster
	than Spark, resulting in a $1.79\times$ to $2.71\times$ speed-up ratio on the makespan.
	The problem we observed was two fold.
	Firstly, the process ID of the workers' threads is not reported correctly and 
	leading to falsely excessive overhead for Spark.
	Secondly, the Spark workers were restarting due to lack of memory, leading 
	to task recomputation and having to reload data into memory.
}

@@

In a follow-up experiment, we further optimized our Kmeans implementations for both Dask and Spark to minimize the size of data kept in memory.
Figure~{\ref{fig:dask_kmeans_gantt}} and Figure~{\ref{fig:spark_kmeans_gantt}} show the Gantt chart for the execution
of this new experiments for Dask and Spark.
The x-axis shows the time in second while the y-axis depicts the workers, with 
one worker thread ID (Dask) and process ID (Spark) per line.
We represent each task on the Gantt chart with its associated function's color.
We again observe that Spark reports three times more processes than Dask threads.
\HL{
	It is unclear what is the cause for reporting extra Spark processes.
	However, with our analysis we found that both Dask and Spark use the same 
	amount of maximum concurrent processes or threads.
	Additionally, the is no significant difference in the average amount of 
	concurrent processes or threads between Dask and Spark.
}

@@

\subsection{Memory management}
% TODO remore reference to figure.
\HL{Our initial Kmeans results showe a significant difference between Dask and Spark.}
The difference in performance \HL{came} from Spark worker restarting due to running out of memory; thus, Spark \HL{needed} to recompute several tasks.
This is unexpected, given that Dask and Spark \HL{were} configured similarly and \HL{were} allocated the same amount of resources.
We think those issues arise from a combination of
(1) the reserved memory for the JVM heap space,
(2) the serialization of Python to Java kept in memory,
(3) Spark using processes compared to Dask using threads, thus limiting shared memory consumption,
and (4) different policies to persist data to disk when running out of memory.

@@

\subsection{Recommendations}
\HL{
	The results from our experiments show that the performance difference between
	Dask and Spark is minimal, in general. 
	The main differences we observed are qualitative.
	We found that Dask integration with the scientific Python ecosystem made it
	easier to implement neuroimaging application.
	Spark configuration required to set multiple environment variables.
	In comparison, Dask is ready-to-go after an installation through PIP.
	Dask dashboard provides more information than Spark ones, making
	the debugging of Dask application easier and faster.
	Therefore, for the neuroimaging field, we recommend the use of Dask over Spark.
	However, in other fields with non-Python-based ecosystem, both Dask and spark
	would be equal candidate as a Big Data engine.
}

\HL{
	% TODO add caveats to avoid when using the engines.

}

\subsection{Limitations}
To make a fair comparison between Spark and Dask, we set both engine configurations
to be equivalent to the best of our capabilities, which comes with some limitations.
Firstly, both engines use a shared implementation for the algorithms.
This is done to keep the theoretical runtime equivalent; however, the chosen
implementation might not be optimal for either of the engines.
Secondly, we decided to use Lustre as a file system for both Spark and Dask.
Spark could benefit from using HDFS as a file system, but HDFS is uncommon in HPC.
\HL{Finally, our Lustre filesystem suffer from an I/O bottleneck coming from 
either the network or disks bandwidth.}
As explained in section~{\ref{subsec:overhead-tradeoff}}, the data transfer 
time is tightly correlated to the overhead of the application.
However, while reducing this bottleneck should improve the performance of 
both engines similarly, it can be hard to predict the actual behavior.

\HL{
	% TODO discuss UCX and other high performance networks.???

}
